---
title: "All Dolphins"
author: "Grace Schotanus, Jonathan VanOyen, and Trevor VanVeldhuisen"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document:
    fig_height: 4
    fig_width: 6
  word_document:
    fig_height: 4
    fig_width: 6
  html_document:
    fig_height: 4
    fig_width: 6
---

```{r, setup, include = FALSE}
# load packages that are going to be used
require(tidyverse)   # this loads mosaic, ggformula, etc. too
require(ggformula)
require(mosaic)
require(glmmTMB)
require(s245)
require(ggeffects)
require(geepack)
require(DHARMa)

# Some customization.  You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw(base_size=12))     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  echo = TRUE,      # for homework, always show R code (this is the default)
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small",   # slightly smaller font for code
  message = FALSE, warning = FALSE) # don't print warnings or messages in compiled document. So you MUST check them in RStudio!
```


```{r}
Dolphin6 <- read.csv('110606-Behavior_QC.csv')
```

```{r}
Dolphin7 <- read.csv('110607-Behavior_QC.csv')
```

```{r}
Dolphin8 <- read.csv('110608-Behavior_QC.csv')
```

```{r}
Dolphin10 <- read.csv('110610-Behavior_QC.csv')
```

```{r}
alldolph <- bind_rows(Dolphin6, Dolphin7, Dolphin8, Dolphin10)
```

```{r}
glimpse(alldolph)
```

```{r}
alldolph = alldolph %>%
  mutate(Ptt = case_when(Ptt == 110606 ~ 'Dolphin6',
Ptt == 110607 ~ 'Dolphin7',
Ptt == 110608 ~ 'Dolphin8',
Ptt == 110610 ~ 'Dolphin10'))
```

```{r}
wide_dolph_dives <- alldolph %>%
  pivot_wider(names_from = What,
              # variables listed in values_from are ones you want to keep/use 
              # that are DIFFERENT for dive and surfacing
              values_from = c(X, Number, Shape, DepthMin, DepthMax, DurationMin, DurationMax,
                              Count, Shallow, Deep, Start, End, start, end, t.diff, flag,
                              depth, duration)
  ) %>%
  # remove variables that are all NA
  janitor::remove_empty(which = 'cols') %>%
  # make datetime variables datetime objects - will be easier for plotting
  mutate(across(Start_Dive:end_Surface, lubridate::mdy_hm))
```

```{r}
cluster_data <- wide_dolph_dives %>%
  select(depth_Dive, duration_Dive) %>%
  mutate(depth=scale(depth_Dive),
         duration=scale(duration_Dive))
```

```{r}
c_out <-cluster::clara(cluster_data, k=2, metric = c("euclidean"))
```


```{r}
wide_dolph_dives <- wide_dolph_dives %>%
  mutate(dtype=c_out$clustering)
```

```{r}
wide_dolph_dives = wide_dolph_dives %>%
mutate(dtype = case_when(dtype == 1 ~ 'shallow',
dtype == 2 ~ 'deep'))
```

Trevor---------------------------------------------
Looking at a summary of all of the variables
```{r}
summary(wide_dolph_dives)
```


Exploratory Graph 
```{r}
gf_point(duration_Surface ~ duration_Dive, data = wide_dolph_dives, color = ~dtype, alpha = 0.2) %>%
  gf_labs( x = 'Dive Duration',
           y = 'Surface Duration') %>%
   gf_theme(scale_color_discrete('Dive Type'))
```
```{r}
gf_point(duration_Surface ~ duration_Dive | ~dtype, data = wide_dolph_dives, alpha = 0.2) %>%
  gf_labs( x = 'Dive Duration',
           y = 'Surface Duration')
```
Fitting the model and summary
```{r}
dolphdive <- glmmTMB(duration_Surface ~ duration_Dive*dtype + depth_Dive + (1| Ptt), data = wide_dolph_dives, family = Gamma(link='log')) 
summary(dolphdive)
```
```{r}
confint(dolphdive)
```

Creating predictors and Residuals

```{r}
conditions <- wide_dolph_dives %>%
  mutate(preds = predict(dolphdive),
         resids = resid(dolphdive))
```
Model Assessment

```{r}
dolphin_dur <- simulateResiduals(dolphdive)
gf_point(dolphin_dur$scaledResiduals ~ fitted(dolphdive),
 alpha = 0.2) %>%
 gf_labs(x = 'Fitted Values',
 y = 'Scaled Residuals')
```
```{r}
s245::gf_acf(~dolphdive) %>%
  gf_lims(y = c(-1,1))
```

Predicted Values of Surface Duration
```{r}
ggeffects::ggpredict(dolphdive)
```




```{r}
d2 <- wide_dolph_dives %>%
  select(duration_Dive, duration_Surface, dtype, depth_Dive) %>%
  na.omit() %>%
  mutate(preds = predict(dolphdive))
```

``{r}
ggpredict(dolphdive, 
          terms = c('duration_Dive',
                    'dtype [0]',
                    'depth_Dive [1]'),
          type = 'fixed') %>%
  plot() %>%
  gf_labs(y = 'Predicted Duration on Surface',
          x = 'Dive Duration')
```






---------------------------------------------------------------------------------------------------------

ALL OF THIS IS DOLPHIN 8:




```{r}
#gf_bar( ~ DepthMax, data = divedata) 
```



```{r}
AllDolphinsBin <- wide_dolph_dives %>%
   mutate(binned_dur = case_when(duration_Dive < 200 ~ "< 200m",
                                 duration_Dive >=200 & duration_Dive < 400 ~ "200 - 400s",
                                 duration_Dive >=400 & duration_Dive < 600 ~ "400 - 600s",
                                 duration_Dive >=600 & duration_Dive < 800 ~ "600 - 800s"))
```

```{r}
gf_boxplot(depth_Dive ~ binned_dur, alpha = 0.2, data = AllDolphinsBin %>%
             select(depth_Dive, binned_dur) %>%
             na.omit()) %>%
  gf_labs(x = 'Binned Duration', y = "Dive Depth")
```



----------

Looking at the distribution of dive depths and duration 

Initial look at depth and duration of a dive:
```{r}
gf_point(depth_Dive ~ duration_Dive,
 data = wide_dolph_dives, alpha = 0.2)%>%
  gf_labs( x = 'Dive Duration',
           y = 'Dive Depth')
```



Depth and Duration of Dives with separation by new variable dtype
```{r}
gf_point(depth_Dive ~ duration_Dive,
 data = wide_dolph_dives,
 color= ~dtype, alpha = 0.2) %>%
  gf_labs( x = 'Dive Duration',
           y = 'Dive Depth')
```
```{r}
wide_dolph_dives <- wide_dolph_dives %>% group_by(Ptt) %>%
  mutate(tsec = as.numeric(Start_Dive - first(Start_Dive)),
         time_block = cut_width(tsec, width = 6*60*60, boundary = 0))
glimpse(wide_dolph_dives)
```

```{r}
dolphin_bin <- glmmTMB(depth_Dive ~ duration_Dive*dtype + duration_Surface + (1|Ptt) + (1|time_block), data = wide_dolph_dives, family = Gamma(link='log')) 

summary(dolphin_bin)
```

ACF graph for model
```{r}
s245::gf_acf(~dolphin_bin) %>%
  gf_lims(y = c(-1,1))
```

Scaled resid. by fitted model 
```{r}
require(DHARMa)
dolphin_sim <- simulateResiduals(dolphin_bin)
gf_point(dolphin_sim$scaledResiduals ~ fitted(dolphin_bin),
 alpha = 0.2) %>%
 gf_labs(x = 'Fitted Values',
 y = 'Scaled Residuals')
```


Predicted Values of Dive Depth
```{r}
ggeffects::ggpredict(dolphin_bin)
```







