---
title: "Stat 245 R/RStudio Intro"
author: "Trevor VanVeldhuisen"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    fig_height: 5
    fig_width: 8
  pdf_document:
    fig_height: 5
    fig_width: 8
  word_document:
    fig_height: 5
    fig_width: 8
---

```{r, setup, include = FALSE}
# load packages that are going to be used
require(tidyverse)   # this loads mosaic, ggformula, etc. too
require(ggformula)
require(mosaic)

# Some customization.  You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw(base_size=12))     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  echo = TRUE,      # for homework, always show R code (this is the default)
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small",   # slightly smaller font for code
  message = FALSE, warning = FALSE) # don't print warnings or messages in compiled document. So you MUST check them in RStudio!
```


<!-- Put your work below here.  Put text in text chunks, code in R chunks. -->

```{r}
globaldata <- readr::read_csv('https://sldr.netlify.app/data/gapminder_clean.csv')
```

```{r}
glimpse(globaldata)
```



```{r}
gf_point(income ~ life_expectancy,
         color = ~four_regions,
         data=globaldata) %>%
  gf_labs(title = "Income and Life Expectancy in the Four Regions",
          x = "Life Expectancy in Years",
          y = "Income") %>%
  gf_theme(scale_fill_discrete('Four Regions'))
  
```



  This graph shows the relationship between income and life expectancy in the four regions. Looking at the general relationship in all four regions, we can tell that the lower the income, the lower the life expectancy. The higher the life expectancy, the higher the income seems to be. Individually, Africa seems to have lower income than the other three regions, however, their life expectancy still increases in some areas without an increase in income. This shows that in Africa life expectancy is not dependent on income as are the other three regions. In the Americas, Asia, and Europe, there is an exponential increase in income as people live longer.





  